Blockchain System

### 2020-12-07 papers/execution-model/model.tex
update
clarified synchronicity, added proposal for type language

\begin{definition}%[Program]
\emph{Expressions} are defined by the following grammar:
\begin{align*}
  \CONSTANT & ::= \ADDR \mid \OPH \mid \PUH \mid\PUK \mid \CODE \mid
              \NTEZ \mid \SUNIT \mid \FALSE \mid \TRUE \\
	\EXPR & :: =  \CONSTANT \mid \VARIABLE \mid \lambda \VARIABLE. \EXPR\ 
	\mid \EXPR\EXPR \mid (\EXPR,\EXPR) \mid \FST\ \EXPR \mid \SND\ \EXPR 	%\mid \NI \EI
	\\&\mid \EXPR\ \PLUS\ \EXPR\ \mid \EXPR\ \MINUS\ \EXPR\ \mid \EXPR\ \EQUAL\ \EXPR\ \mid \EXPR\ \LESS\ \EXPR
	\\&\mid \EXPR\ \AND\ \EXPR\ \mid \EXPR\ \OR\ \EXPR\ \mid \NOT\ \EXPR
  \\&\mid \NIL\ \mid \CONS\ \EXPR\ \EXPR \mid \LEFT\ \EXPR \mid\RIGHT\
  \EXPR \mid \SOME\ \EXPR \mid \NONE
  \\& \mid \RAISE\ \EXPR \mid \TRY\ \EXPR\ \EXCEPT\ \EXPR
  \\& \mid \MATCH\ \EXPR\ \WITH\ \PATTERN\to\EXPR \dots
  \\& \mid \TRANSFER[\EXPR]\EXPR\EXPR\EXPR\EXPR
  \\& \mid \ORIGINATE\EXPR\EXPR\EXPR\EXPR\EXPR
  \\&\mid \QOP\  \EXPR\\
  \QOP & ::= balance \mid status \mid storage \mid code \mid counter
\end{align*}
\end{definition}

@@

\begin{mathpar}
  \inferrule[Node-Opp]
    {
    \Angle{\PAK,\OP.\PUK} \in \ACCOUNTS
  }{
    [\EC\OP :: \EXPRS, \ACCOUNTS, \OPERATIONS] \NodeTrans
    [\EXPRS, \ACCOUNTS, (\EC{},\OP):: \OPERATIONS]
  }
\end{mathpar}

\begin{mathpar}
  \inferrule{
    \CHECKBAL (\MANAGERS, \PUK, \NTEZ, \MTEZ) \\
    \CHECKCOU (\MANAGERS, \PUK) \\
    \CHECKPUB (\MANAGERS, \PUK') \\
    \CHECKGAS (\PUK, \PUK', \NTEZ, \MTEZ) \\
    \OP = \TRANSFER\NTEZ\PUK{\PUK'}\MTEZ \\
    \OPH = \GENERATEOPH (\PUK, \PUK', \NTEZ, \MTEZ, \TIME)
  }{
   { [\EXPRS, \ACCOUNTS, (\EC{},\OP) :: \OPERATIONS] \| [\PENDING,
     \ACCEPTED, \MANAGERS, \CONTRACTORS, \TIME] } \SystemTrans \\
   { [\EC{?\OPH}  :: \EXPRS, \ACCOUNTS, \OPERATIONS] \|
     [ \Angle{\OP,\OPH, \TIME}
     ::\PENDING,
     \ACCEPTED,
     \UPDATECOU(\MANAGERS, \PUK, \TRUE),
     \CONTRACTORS,
     \TIME]
   }
  }
\end{mathpar}

@@

\begin{mathpar}
  \inferrule[Node-Op]
  {
    \Angle{\PAK,\OP.\PUK} \in \ACCOUNTS
  }{
    [\EXPRS, \ACCOUNTS, \OPERATIONS]
    \NodeTrans
    [\EXPRS, \ACCOUNTS, \OP:: \OPERATIONS]
  }

  \inferrule[Node-Oph]
  {
   \\
  }{
    [\EC\OP :: \EXPRS, \ACCOUNTS, \OPERATIONS]
    \NodeTrans
    [\EC\OPH :: \EXPRS, \ACCOUNTS, \OPERATIONS]
  }

>>>>

\begin{definition}%[Program]
\emph{Expressions} are defined by the following grammar:
\begin{align*}
  \CONSTANT & ::= \ADDR \mid \OPH \mid \PUH \mid\PUK \mid \CODE \mid
              \NTEZ \mid \SUNIT \mid \FALSE \mid \TRUE \\
  \STATUS & ::= \STATUSPENDING \mid \STATUSINCLUDING (\dots) \mid \STATUSTIMEOUT \\
	\EXPR & :: =  \CONSTANT \mid \STATUS \mid \VARIABLE \mid \lambda \VARIABLE. \EXPR\ 
	\mid \EXPR\EXPR \mid (\EXPR,\EXPR) \mid \FST\ \EXPR \mid \SND\ \EXPR 	%\mid \NI \EI
	\\&\mid \EXPR\ \PLUS\ \EXPR\ \mid \EXPR\ \MINUS\ \EXPR\ \mid \EXPR\ \EQUAL\ \EXPR\ \mid \EXPR\ \LESS\ \EXPR
	\\&\mid \EXPR\ \AND\ \EXPR\ \mid \EXPR\ \OR\ \EXPR\ \mid \NOT\ \EXPR
  \\&\mid \NIL\ \mid \CONS\ \EXPR\ \EXPR \mid \LEFT\ \EXPR \mid\RIGHT\
  \EXPR \mid \SOME\ \EXPR \mid \NONE
  \\& \mid \RAISE\ \EXPR \mid \TRY\ \EXPR\ \EXCEPT\ \EXPR
  \\& \mid \MATCH\ \EXPR\ \WITH\ \PATTERN\to\EXPR \dots
  \\& \mid \TRANSFER[\EXPR]\EXPR\EXPR\EXPR\EXPR
  \\& \mid \ORIGINATE\EXPR\EXPR\EXPR\EXPR\EXPR
  \\&\mid \QOP\  \EXPR\\
  \QOP & ::= balance \mid status \mid storage \mid code \mid counter
\end{align*}
\end{definition}

Example code
\begin{align*}
  \OPH &=  \TRANSFER[\EXPR]\EXPR\EXPR\EXPR\EXPR \\
       & \MATCH\ {status\ \OPH}\ \WITH \\
       & \mid \STATUSPENDING \to \textrm{not yet included, continue waiting} \\
       & \mid \STATUSINCLUDING (n) \to \textrm{included with $n$
         blocks confirmation}\\
       & \mid \STATUSTIMEOUT \to \textrm{lost, try again}
\end{align*}
Grammar of types
\begin{align*}
  \TYPE & ::= \ADDR \mid \OPH \mid \PUH \mid \PUK \mid \CODE\ \TYPE \mid
          \TINT \mid \TUNIT \mid \TBOOL \mid \TSTRING \mid \TSTATUS \\
  & \mid \TYPE\to\TYPE \mid \TPAIR\ \TYPE\ \TYPE \mid \TLIST\ \TYPE
    \mid \TSUM\ \TYPE\ \TYPE \mid \TOPTION\ \TYPE \mid \TEXCEPTION
\end{align*}

@@

\begin{mathpar}
  \inferrule[Node-Eval]
  {
    \EXPR \ExprTrans \EXPR'
  }{
    [\EC\EXPR :: \EXPRS, \ACCOUNTS, \OPERATIONS] \NodeTrans{}
    [\EC{\EXPR'} :: \EXPRS, \ACCOUNTS, \OPERATIONS]
  }
\end{mathpar}

\begin{mathpar}
  \inferrule[Node-Op']
    {
    \Angle{\PAK,\OP.\PUK} \in \ACCOUNTS
  }{
    [\EC\OP :: \EXPRS, \ACCOUNTS, \OPERATIONS] \NodeTrans{}
    [\EXPRS, \ACCOUNTS, (\EC{},\OP):: \OPERATIONS]
  }
\end{mathpar}

\begin{mathpar}
  \inferrule[Node-Inject']{
    \CHECKBAL (\MANAGERS, \PUK, \NTEZ, \MTEZ) \\
    \CHECKCOU (\MANAGERS, \PUK) \\
    \CHECKPUB (\MANAGERS, \PUK') \\
    \CHECKGAS (\PUK, \PUK', \NTEZ, \MTEZ) \\
    \OP = \TRANSFER\NTEZ\PUK{\PUK'}\MTEZ \\
    \OPH = \GENERATEOPH (\PUK, \PUK', \NTEZ, \MTEZ, \TIME)
  }{
   { [\EXPRS, \ACCOUNTS, (\EC{},\OP) :: \OPERATIONS] \| [\PENDING,
     \ACCEPTED, \MANAGERS, \CONTRACTORS, \TIME] } \SystemTrans \\
   { [\EC{\OPH}  :: \EXPRS, \ACCOUNTS, \OPERATIONS] \|
     [ \Angle{\OP,\OPH, \TIME}
     ::\PENDING,
     \ACCEPTED,
     \UPDATECOU(\MANAGERS, \PUK, \TRUE),
     \CONTRACTORS,
     \TIME]
   }
  }
\end{mathpar}

\begin{mathpar}
  \inferrule[Block-Accept'']{
    \OP = \TRANSFER\NTEZ\PUK{\PUK'}\MTEZ \\
    \TIME'-\TIME < 60
  }{
    { % [\EC{?\OPH}  :: \EXPRS, \ACCOUNTS, \OPERATIONS] \|
      [\Angle{\OP,\OPH, \TIME} :: \PENDING, \ACCEPTED, \MANAGERS,
      \CONTRACTORS, \TIME']}
    \BlockTrans  \\
    { % [\EC{\OPH}  :: \EXPRS, \ACCOUNTS, \OPERATIONS] \|
      [\PENDING, \Angle{\OP,\OPH, \TIME,\TIME'} :: \ACCEPTED,
      \UPDATESUCC(\MANAGERS, \PUK, \PUK', \NTEZ, \MTEZ),
      \CONTRACTORS, \TIME'+1]}
  }
\end{mathpar}

\begin{mathpar}
  \inferrule[Block-Timeout'']{
    \OP = \TRANSFER[\STRING]\NTEZ\PUK\ADDR\MTEZ \\
    \TIME'-\TIME \ge 60
  }{ % [\EC{?\OPH}  :: \EXPRS, \ACCOUNTS, \OPERATIONS] \|
    {[\Angle{\OP,\OPH, \TIME} :: \PENDING, \ACCEPTED, \MANAGERS,
      \CONTRACTORS, \TIME']}
    \BlockTrans 
    { % [\EC{\RAISE timeout}  :: \EXPRS, \ACCOUNTS, \OPERATIONS] \|
      [\PENDING, \ACCEPTED, \UPDATECOU(\MANAGERS, \PUK, \FALSE),
      \CONTRACTORS, \TIME']}
  }
\end{mathpar}

\begin{itemize}
\item library call \texttt{call\_contract} with \texttt{confirmations:None} corresponds to
  \begin{enumerate}
  \item{} first \TirName{Node-Op'}, \TirName{Node-Inject'}
  \item{} first \TirName{Node-Op'}, \TirName{Node-Inject'}
  \item{} first \TirName{Node-Op'}, \TirName{Node-Reject(balance,count,pubkey,fee)}
  \end{enumerate}
\item afterwards must check the status to find whether the
  transfer is still pending or \TirName{Block-Accept''} or
  \TirName{Block-Timeout''} has happened.
\item no further transfer is possible before the status is resolved to
  accept or timeout.
\end{itemize}

@@

\begin{mathpar}
  \inferrule[Node-Op]
  {
    \Angle{\PAK,\OP.\PUK} \in \ACCOUNTS
  }{
    [\EXPRS, \ACCOUNTS, \OPERATIONS]
    \NodeTrans{}
    [\EXPRS, \ACCOUNTS, \OP:: \OPERATIONS]
  }

  \inferrule[Node-Oph]
  {
   \\
  }{
    [\EC\OP :: \EXPRS, \ACCOUNTS, \OPERATIONS]
    \NodeTrans{}
    [\EC\OPH :: \EXPRS, \ACCOUNTS, \OPERATIONS]
  }
